Enriching Object Detection with 2D-3D Registration and Continuous Viewpoint Estimation

### 2015-04-07 2349_ext.tex
modified ext abstract

A large body of recent work on object detection has focused on providing more
than a bounding box of an object. Some exploited 3D CAD
model databases to provide viewpoint estimation. These approaches work
by aligning exact 3D models to images using templates generated from renderings
of the 3D models at a set of discrete viewpoints.

@@

The result is a whitened feature representation, termed WHO (Whitened Histogram
of Orientations). This development makes it feasible to train hundreds of
thousands of mid-level patch detectors for recognition.

Though it achieved remarkable improvements, generating LDA templates (matrix
decomposition with Gaussian Elimination) and calibrating them are
computationally expensive and require gigabytes of memory and storage, while
the viewpoint discretization hampers pose estimation performance.

@@

\begin{figure}[t]
  \centering
  % \includegraphics[width=0.9\linewidth]{schematics} 
  \includegraphics[width=0.8\linewidth]{front} % \\[-5pt]
  \caption{Using a database of 3D CAD models, we generate NZ-WHO
    templates which can be used to either detect objects directly or
    enrich the output of an existing detector with high-quality,
    continuous pose and 3D CAD model exemplar.}
    % initialize pose of an object in the bounding box from other
    % detectors. Once initialization is given, we use NZ-WHO to propose
    % and validate plausible pose on-the-fly and
    % further tune the translation, 3D viewpoint and focal length continuously.}
  % \vspace{-1em}
  \label{fig:front}
\end{figure}

\paragraph{Overview}

In this paper, we propose a novel method for 2D-3D alignment of exemplar CAD
models to real-world images that circumvents the need for calibration and
greatly enhances the scalability of WHO. As a result, we can render novel views
and train corresponding exemplar models {\em on-the-fly}, without the need for
offline processing. We call these Non-Zero Whitened Histograms of Orientations
(NZ-WHO) templates Fig.~\ref{fig:front}. As a by-product, we can formulate the
alignment problem as a parameter search in a continuous pose space, consisting
of yaw, pitch and roll, which we implement using MCMC sampling.

@@

Second, we demonstrate that our method can enrich the output of an existing
object class detector, such as the DPM or the
R-CNN with additional 3D information. By applying our
method to candidate detections provided by the respective detector, we can
augment the original detections with an estimate of 3D continuous pose and a 3D
CAD model exemplar.

Finally, we give an in-depth experimental study that demonstrates
the effectiveness of our approach on a standard benchmark for
object detection and viewpoint estimation~\cite{Xiang14}.

\begin{figure}[t]
\small
\setlength\tabcolsep{1pt}
\centering
\begin{tabular}{ccccc}
  \includegraphics[width=0.2\textwidth]{car_cnn/2c.png} &
  \includegraphics[width=0.2\textwidth]{car_cnn/2e.png} \\
  \includegraphics[width=0.2\textwidth]{bicycle_cnn/4b.png} &
  \includegraphics[width=0.2\textwidth]{bicycle_cnn/4c.png} \\
\end{tabular}
\caption{Example enriched bounding boxes. Given R-CNN
	detection bounding boxes, our method predicts the best 2D-3D
	registration given the bounding box input. Blue boxes are R-CNN output and
	purple boxes are the tightest bounding box enclosing predicted CAD model.}
\label{fig:pascal12cnn}
\end{figure}

\paragraph{Non-Zero Whitened Histograms of Orientations}

@@

\paragraph{Fine-Tuning using MCMC}
Due to the speed of our template synthesis procedure, we are able to perform
joint optimization of scale, translation, continuous rotation, and focal length
using the Metropolis-Hastings algorithm.

@@

\paragraph{Experiments}

>>>>

A large body of recent work on object detection has focused on providing more
information than a bounding box of an object. Some used 3D CAD
model databases to provide viewpoint estimation. These approaches work
by aligning exact 3D models to images using templates generated from renderings
of the 3D models at a set of discrete viewpoints.

@@

The result is a whitened feature representation, termed WHO (Whitened Histogram
of Orientations). This development makes it feasible to train a large number of
mid-level patch detectors for recognition.

Though generating and calibrating LDA templates (matrix decomposition with
Gaussian Elimination) achieved remarkable improvements, they are
computationally expensive and use a prohibitive amount of memory and storage.
In addition, viewpoint discretization hampers pose estimation performance.

@@

\begin{figure}[t]
  \centering
  % \includegraphics[width=0.9\linewidth]{schematics} 
  \includegraphics[width=0.9\linewidth]{front} % \\[-5pt]
  \caption{Using a database of 3D CAD models, we generate NZ-WHO
    templates which can be used to either detect objects directly or
    enrich the output of an existing detector with high-quality,
    continuous pose and 3D CAD model exemplar.}
    % initialize pose of an object in the bounding box from other
    % detectors. Once initialization is given, we use NZ-WHO to propose
    % and validate plausible pose on-the-fly and
    % further tune the translation, 3D viewpoint and focal length continuously.}
  % \vspace{-1em}
  \label{fig:front}
\end{figure}

\vspace{-0.1in}
\paragraph{Overview}

In this paper, we propose a novel method for 2D-3D alignment of exemplar CAD
models to real-world images that circumvents the need for calibration and
greatly enhances the scalability of WHO. As a result, we can render novel views
and train corresponding exemplar models {\em on-the-fly}, without the need for
offline processing. We call these Non-Zero Whitened Histograms of Orientations
(NZ-WHO) templates. First, we generate an ensemble of NZ-WHO templates to find
coarse viewpoints and fine-tune them using MCMC. The schematic of our pipeline
is on Fig.~\ref{fig:front}. As a by-product, we can formulate the alignment
problem as a parameter search in a continuous pose space, consisting of yaw,
pitch and roll, which we implement using MCMC sampling.

@@

Second, we demonstrate that our method can enrich the output of an existing
object class detector, such as the DPM or the
R-CNN with additional 3D information. By applying our
method to candidate detections provided by the respective detector, we augment the original detections with an estimate of 3D continuous pose and a 3D
CAD model exemplar.

Finally, our in-depth experimental study demonstrates
the effectiveness of the approach on a standard benchmark for
object detection and viewpoint estimation~\cite{Xiang14}.

\begin{figure}[t]
\small
\setlength\tabcolsep{1pt}
\centering
\begin{tabular}{ccccc}
  \includegraphics[width=0.18\textwidth]{car_cnn/2c.png} &
  \includegraphics[width=0.18\textwidth]{car_cnn/2e.png} \\
  \includegraphics[width=0.18\textwidth]{bicycle_cnn/4b.png} &
  \includegraphics[width=0.18\textwidth]{bicycle_cnn/4c.png} \\
\end{tabular}
\caption{Example enriched bounding boxes. Given R-CNN
	detection bounding boxes, our method predicts the best 2D-3D
	registration given the bounding box input. Blue boxes are R-CNN output and
	purple boxes are the tightest bounding box enclosing predicted CAD model.}
\label{fig:pascal12cnn}
\end{figure}

\vspace{-0.1in}
\paragraph{Non-Zero Whitened Histograms of Orientations}

@@

\vspace{-0.1in}
\paragraph{Fine-Tuning using MCMC}
Due to the speed of our template synthesis procedure, we are able to perform
joint optimization of scale, translation, continuous rotation, and focal length
using the Metropolis-Hastings algorithm.

@@

\vspace{-0.1in}
\paragraph{Experiments}
